---
title: "Tandem repeats analysis"
author: "Pavel Nikitin, Sviatoslav Sidorov"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

Load libraries:

```{r}
library(dplyr)
library(ggplot2)
library(tidyr)
library(stringr)
stringsAsFactors = F
```

Parse arguments:

```{r}
args = commandArgs()
contigs.N.path = args[9]
contigs.V.path = args[10]
units.N.path = args[11]
units.V.path = args[12]
```

Upload tables of contigs and repeat units:

```{r}
contigs.n <- read.delim(contigs.N.path, header = F,  sep="\t", stringsAsFactors = F)
names(contigs.n) = c("contig", "length", "mean.coverage", "repr", "unit.seq")
contigs.n <- contigs.n %>% drop_na()

contigs.v <- read.delim(contigs.V.path, header = F,  sep="\t", stringsAsFactors = F) 
names(contigs.v) = c("contig", "length", "mean.coverage", "repr", "unit.seq")
contigs.v <- contigs.v %>% drop_na()

all.repeats.n <- read.delim(units.N.path, header = F,  sep="\t", stringsAsFactors = F) 
names(all.repeats.n) = c("contig", "length", "mean.coverage", "repr", "unit.seq")

all.repeats.v <- read.delim(units.V.path, header = F,  sep="\t", stringsAsFactors = F)
names(all.repeats.v) = c("contig", "length", "mean.coverage", "repr", "unit.seq")
```

Define a function to calcualte GC-content:

```{r}
get_GC_content <- function(unit_seq) {
  num_g <- str_count(unit_seq, "G")
  num_c <- str_count(unit_seq, "C")
  gc_content <- (num_g + num_c) / str_length(unit_seq) * 100 
  return(as.numeric(gc_content))
}
```

Calculate the GC-content and length of contigs and repeat units:

```{r}
contigs.n <- contigs.n %>%
  mutate(GC = get_GC_content(unit.seq)) %>%
  mutate(unit.length = nchar(unit.seq)) 
contigs.v <- contigs.v %>%
  mutate(GC = get_GC_content(unit.seq)) %>%
  mutate(unit.length = nchar(unit.seq)) 

all.repeats.n <- all.repeats.n %>%
  mutate(GC = get_GC_content(unit.seq)) %>%
  mutate(unit.length = nchar(unit.seq)) 
all.repeats.v <- all.repeats.v %>%
  mutate(GC = get_GC_content(unit.seq)) %>%
  mutate(unit.length = nchar(unit.seq))
```

Create an extended table for contigs to normalize on the read coverage:

```{r}
contigs.n.subset = contigs.n %>%
  dplyr::select(mean.coverage, GC, unit.length) 

contigs.n.extended = as.data.frame(lapply(contigs.n.subset, rep, contigs.n.subset$mean.coverage)) %>%
  mutate(species = "D. raddei nairensis")

contigs.v.subset = contigs.v %>%
  dplyr::select(mean.coverage, GC, unit.length) 

contigs.v.extended = as.data.frame(lapply(contigs.v.subset, rep, contigs.v.subset$mean.coverage)) %>%
  mutate(species = "D. valentini")

contigs.extended = dplyr::bind_rows(contigs.n.extended, contigs.v.extended) %>% 
  mutate(species = factor(species, levels = c("D. raddei nairensis", "D. valentini")))
```

Create an extended table for repeat units to normalize on the read coverage:

```{r}
all.repeats.n.subset = all.repeats.n %>%
  dplyr::select(mean.coverage, GC, unit.length) 

all.repeats.n.extended = as.data.frame(lapply(all.repeats.n.subset, rep, all.repeats.n.subset$mean.coverage)) %>%
  mutate(species = "D. raddei nairensis")

all.repeats.v.subset = all.repeats.v %>%
  dplyr::select(mean.coverage, GC, unit.length) 

all.repeats.v.extended = as.data.frame(lapply(all.repeats.v.subset, rep, all.repeats.v.subset$mean.coverage)) %>%
  mutate(species = "D. valentini")

all.repeats.extended = dplyr::bind_rows(all.repeats.n.extended, all.repeats.v.extended) %>% 
  mutate(species = factor(species, levels = c("D. raddei nairensis", "D. valentini")))
```

Plot GC-content distribution of contigs:

```{r}
plot.contigs.gc = ggplot() +
  geom_histogram(data = contigs.n.extended,
                 alpha = 1,
                 aes(x = GC,
                     y = stat(count / sum(count)),
                     fill = "D. raddei nairensis"),
                 bins = 100) +
  geom_histogram(data = contigs.v.extended,
                 alpha = 0.7,
                 aes(x = GC,
                     y = stat(count / sum(count)),
                     fill = "D. valentini"),
                 bins = 100) +
  scale_fill_manual(values = c("D. raddei nairensis" = "#da2c38",
                               "D. valentini" = "#00d103"),
                    name = "Species") +
  xlab("Contig GC-content") +
  ylab("Normalized sum coverage") +
  scale_x_continuous(breaks = seq(0, 100, 5),
                     limits = c(-5, 105)) +
  scale_y_continuous(breaks = seq(0, 0.08, 0.01),
                     limits = c(-0.001, 0.08)) +
  theme_classic() +
  theme(legend.position = c(.95, .70),
        legend.justification = c("right", "top"),
        text = element_text(size = 13))

ggsave("contig_gc_content.pdf",
       plot = plot.contigs.gc,
       width = 15,
       height = 10,
       scale = 1, 
       units = c("cm"),
       limitsize = FALSE)
```

Plot GC-content distribution of repeat units:

```{r}
plot.repeats.gc = ggplot() +
  geom_histogram(data = all.repeats.n.extended,
                 alpha = 1,
                 aes(x = GC, y = stat(count / sum(count)),
                     fill = "D. raddei nairensis"),
                 binwidth = 5) +
  geom_histogram(data = all.repeats.v.extended,
                 alpha = 0.7,
                 aes(x = GC, y = stat(count / sum(count)),
                     fill = "D. valentini"),
                 binwidth = 5) +
  scale_fill_manual(values = c("D. raddei nairensis" = "#da2c38",
                               "D. valentini" = "#00d103"),
                    name = "Species") +
  xlab("Unit GC-content") +
  ylab("Normalized sum coverage") +
  scale_x_continuous(breaks = seq(0, 100, 5))+
  scale_y_continuous(breaks = seq(0, 0.3, 0.1),
                     limits = c(0, 0.3)) +
  theme_classic() +
  theme(legend.position = c(.37, .90),
        legend.justification = c("right", "top"),
        text = element_text(size = 15))

ggsave("units_gc_content.pdf",
       plot = plot.repeats.gc,
       width = 15,
       height = 10,
       scale = 1,
       units = c("cm"),
       limitsize = FALSE)
```

Plot sequence length distribution of contigs:

```{r}
contigs.extended %<>% 
  group_by(species) %>%
  mutate(median.length = median(as.numeric(unit.length)))
plot.contigs.length = ggplot(data = contigs.extended,
                             aes(x = unit.length,
                                 fill = species)) +
  geom_histogram(binwidth = 10) +
  geom_vline(aes(xintercept= median.length), colour='black') +
  facet_wrap(~species, ncol = 1) +
  scale_fill_manual(values = c("#da2c38",
                             alpha("#00d103", 0.7))) +
  xlab("Contig length") +
  ylab("Count") +
  theme_classic() +
  theme(legend.position = 'none',
        text = element_text(size = 15))

ggsave("contigs_length.pdf",
       plot = plot.contigs.length,
       width = 15,
       height = 10,
       scale = 1,
       units = c("cm"),
       limitsize = FALSE)
```

Plot sequence length distribution of repeat units:

```{r}
plot.repeats.length = ggplot() +
  geom_histogram(data = all.repeats.n.extended,
                 alpha = 1,
                 aes(x = unit.length, y = stat(count / sum(count)),
                     fill = "D. raddei nairensis"),
                 binwidth = 10) +
  geom_histogram(data = all.repeats.v.extended,
                 alpha = 0.7,
                 aes(x = unit.length, y = stat(count / sum(count)),
                     fill = "D. valentini"),
                 binwidth = 10) +
  scale_fill_manual(values = c("D. raddei nairensis" = "#da2c38",
                               "D. valentini" = "#00d103"),
                    name = "Species") +
  xlab("Unit length") +
  ylab("Normalized sum coverage") +
  scale_x_continuous(breaks = seq(0, 180, 10)) +
  scale_y_continuous(breaks = seq(0, 0.8, 0.1),
                     limits = c(0, 0.8)) +
  theme_classic() +
  theme(legend.position = c(.95, .70),
        legend.justification = c("right", "top"),
        text = element_text(size = 13))

ggsave("units_length.pdf",
       plot = plot.repeats.length,
       width = 15,
       height = 10,
       scale = 1,
       units = c("cm"),
       limitsize = FALSE)
```
